\propertitle

### 2019-04-19 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-19T15:23:02+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy principle was proposed by Jaynes
\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
premisses at a more direct level, as in the approach just described. This
point is made especially clear in \textcite{jaynes1994_r2003},
\sect~11.1:\vspace{0.5\topsep}
\begin{quote}\footnotesize
  If we knew the numbers $N$, $L$, $W$, then this could be solved by direct
  application of Bayes' theorem; without that information, we could still
  introduce the unknowns $N$, $L$, $W$ as nuisance parameters and use
  Bayes' theorem, eliminating them at the end. \textelp{} However, the
  Bayesian solution would not really address our problem; it only transfers
  it to the problem of assigning priors to $N$, $L$, $W$, leaving us back
  in essentially the same situation; how do we assign informative
  probabilities?
\end{quote}
Its idea is to translate \emph{all} data we're given as constraints for our
probability assignment, resolving the leftover freedom by choosing the
probability distribution having highest Shannon entropy, as a measure of
its uncertainty or broadness. It can be shown that the use of this
principle is approximately equivalent, under specific conditions, to a
probability assignment at a deeper level based on an assumption of
exchangeability together with a multinomial probability distribution for
long-run frequencies, or with an assumption about sufficient statistics
\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2009,portamana2017}.

Owing to its directness, the maximum-entropy principle is straightforward
to use -- and to misuse. In our previous example, for instance, it gives
silly results unless the number of time bins is much larger than the
possible activity values (it has thus been variously misused in the
neuroscientific literature \mynote{refs, especially to Tkacik}).

In the present work we show how to use the maximum-entropy principle to...

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-19T19:09:13+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy principle was proposed by Jaynes
\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
premisses at a more direct level, as in the approach just described. This
point is made especially clear in \textcite{jaynes1994_r2003},
\sect~11.1:\vspace{0.5\topsep}
\begin{quote}\footnotesize
  If we knew the numbers $N$, $L$, $W$, then this could be solved by direct
  application of Bayes' theorem; without that information, we could still
  introduce the unknowns $N$, $L$, $W$ as nuisance parameters and use
  Bayes' theorem, eliminating them at the end. \textelp{} However, the
  Bayesian solution would not really address our problem; it only transfers
  it to the problem of assigning priors to $N$, $L$, $W$, leaving us back
  in essentially the same situation; how do we assign informative
  probabilities?
\end{quote}
The idea behind this principle is to translate \emph{all} the data we're
given into constraints -- such as expectations -- for our probability
assignment, resolving the leftover freedom by choosing the constrained
probability distribution having highest Shannon entropy, as a measure of
its uncertainty or broadness. It can be shown that the use of this
principle is approximately equivalent, under specific conditions, to a
probability assignment at a deeper level based on an assumption of
exchangeability together with a multinomial probability distribution for
long-run frequencies, or with an assumption about sufficient statistics
\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2017,portamana2009}.

Sometimes this principle is applied with only \emph{part} of the available
data used as constraint; such an altered approach can be a heuristic way of
testing for sufficient statistics.

Owing to its directness, the maximum-entropy principle is straightforward
to use -- and to misuse. In our previous example, for instance, it would
give silly results if the number of time bins were not much larger than the
possible activity values (it has thus been variously misused in the
neuroscientific literature \mynote{refs, especially to Tkacik}).

In the present work we show how to use the maximum-entropy principle to
assign a probability distribution for the activity of a population of
neurons at a new time bin, given specific statistics about the activity of
a \emph{subpopulation}, that is, a sample, of the full population. This
kind of probability assignment is extremely important in neuroscience: we
try to understand the workings of brain areas that may include tens of
thousands of neurons, by analysing recording of only a few tens or hundreds
of them.
